manuscript_modelling_and_plots_takeovers

### 2024-08-15 manuscript_and_modelling_plots_takeovers.Rmd
initialing analysis script and data for the analysis

Tidy up longitudinal jerk modelling

Model lateral jerk to see if model fit and inferences are still good (with maximal random effects)

Add random interactions to all the other models

Adjust the plots and tables

Send to co-authors and start on supplemental material.

```{r}
# distribution of negative longitudinal jerk
ggplot(response.vehicle.eye %>%
         dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, ho.rt > .15, collision == 0) %>%
         dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x)) %>%
         dplyr::mutate(ppid = as.factor(ppid)), mapping = aes(max_long_jerk)) +
  geom_histogram() +
  facet_grid(n_back ~ ttc_criticality_factor)

@@

max_long_mixture_weak_priors  <- readRDS(here::here("ITS/Work package 6.5 (HiDrive)/TME study/Analysis/max_long_mixture_weak_prior.rds"))

@@

pp_check(max_long_mixture_6, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
pp_check(max_long_mixture_6, type = "stat_grouped", stat = "mean", group = "n_back")

pp_check(max_long_mixture_6, type = "dens_overlay_grouped", group = "n_back", ndraws = 100)
pp_check(max_long_mixture_6, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

grand_mean_n_back_dist <- max_long_mixture_2_ttc %>% 
  epred_draws(newdata = expand_grid(ttc_criticality_factor = c(as.factor(3), as.factor(5))),
              re_formula = NA)

@@

# model of maximum longitudinal acceleration with hella weak priors. To check if the model is blowing up cause if lack of priors constraining parameter values
max_long_jerk_skewnormal_reg_priors_maximal <- brm(data = response.vehicle.eye %>%
                                             dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, ho.rt > .15, collision == 0) %>%
                                             dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x)) %>%
                                             dplyr::mutate(ppid = as.factor(ppid)),
                                           family = skew_normal(),
                                           prior = c(prior(student_t(3, 0, 2.5), class = "Intercept"),
                                                     prior(student_t(3, 0, 2.5), class = "b")),
                                           max_long_jerk ~ n_back.x * ttc_criticality_factor + (n_back.x * ttc_criticality_factor | ppid),
                                           control = list(adapt_delta = 0.99999999999),
                                           iter = 5000, warmup = 2000, chains = 2, cores = 2, seed = 13,
                                           file = "max_long_jerk_skewnormal_reg_priors_maximal")

print(summary(max_long_jerk_skewnormal_reg_priors_maximal), digits = 5)
describe_posterior(max_long_jerk_skewnormal_reg_priors_maximal, centrality = "mean")

pp_check(max_long_jerk_skewnormal_reg_priors_maximal, type = "dens_overlay_grouped", group = "n_back.x", ndraws = 100)
pp_check(max_long_jerk_skewnormal_reg_priors_maximal, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

@@

# prior for beta_C i.e., the change in max long when TTC = 5 s
prior(uniform(10, 11)) %>% 
  parse_dist() %>% 
  ggplot(aes(xdist = .dist_obj, y = prior)) + 
  stat_halfeye(.width = c(.5, .99)) +
  scale_y_discrete(NULL, breaks = NULL, expand = expansion(add = 0.1)) +
  scale_x_continuous(expression(italic(p)(beta[1]))) +
  theme_bw()

# regularising priors for the rest of the parameters
prior(normal(9, 1)) %>% 
  parse_dist() %>% 
  ggplot(aes(xdist = .dist_obj, y = prior)) + 
  stat_halfeye(.width = c(.5, .99)) +
  scale_y_discrete(NULL, breaks = NULL, expand = expansion(add = 0.1)) +
  scale_x_continuous(expression(italic(p)(beta[1]))) +
  theme_bw()

prior(student_t(30, 9, 2.5)) %>% 
  parse_dist() %>% 
  ggplot(aes(xdist = .dist_obj, y = prior)) + 
  stat_halfeye(.width = c(.5, .99)) +
  scale_y_discrete(NULL, breaks = NULL, expand = expansion(add = 0.1)) +
  scale_x_continuous(expression(italic(p)(beta[1]))) +
  theme_bw()

@@

```{r}
ggplot(response.vehicle.eye %>%
         dplyr::filter(rt.outlier == FALSE, ho.rt != 0, collision == 0), mapping = aes(x = max_lat)) +
  geom_histogram(aes(fill = as.factor(n_back)), alpha = 0.5, bins = 50, col = "black") +
  facet_wrap(~ttc_criticality.x)

>>>>

Calculate the average absolute lateral accelerations were and see if it correlates with maximum lateral accelerations

Find studies that talk about max decelerations being around 12 m/s^2 and how this makes modelling the response difficult by just taking the max value, but averaging over the all the decleration produces some nice fits - I think it's Gold et al or maybe Radlymayr et al who mentions the max longitudinal metric

Further investigate variance in the effect of N-back on mean deceleration.

Check if a Gaussian produces a better fit than the skew gaussian.

```{r}
ggplot(response.vehicle.eye %>%
         dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, ho.rt > .15, collision == 0) %>%
         dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x)) %>%
         dplyr::mutate(ppid = as.factor(ppid)), mapping = aes(mean_decel)) +
  geom_histogram() +
  facet_grid(n_back ~ ttc_criticality_factor)

# regularising priors for parameter values
prior(student_t(5, 9, 2.5)) %>% 
  parse_dist() %>% 
  ggplot(aes(xdist = .dist_obj, y = prior)) + 
  stat_halfeye(.width = c(.5, .99)) +
  scale_y_discrete(NULL, breaks = NULL, expand = expansion(add = 0.1)) +
  scale_x_continuous(expression(italic(p)(beta[1]))) +
  theme_bw()

prior(normal(9, 2)) %>% 
  parse_dist() %>% 
  ggplot(aes(xdist = .dist_obj, y = prior)) + 
  stat_halfeye(.width = c(.5, .99)) +
  scale_y_discrete(NULL, breaks = NULL, expand = expansion(add = 0.1)) +
  scale_x_continuous(expression(italic(p)(beta[1]))) +
  theme_bw()

mean_decel_skewnormal_reg_priors_maximal <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/mean_decel_skewnormal_reg_priors_maximal.rds"))

# model of maximum longitudinal jerk with regularising priors. Random effects structure is maximal
mean_decel_skewnormal_reg_priors_maximal <- brm(data = response.vehicle.eye %>%
                                             dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, ho.rt > .15, collision == 0) %>%
                                             dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x)) %>%
                                             dplyr::mutate(ppid = as.factor(ppid)),
                                           family = skew_normal(),
                                           prior = c(prior(student_t(3, 0, 2.5), class = "Intercept"),
                                                     prior(student_t(3, 0, 2.5), class = "b")),
                                           bf(mean_decel ~ n_back * ttc_criticality_factor + (n_back * ttc_criticality_factor | ppid)),
                                           control = list(adapt_delta = 0.99999999999),
                                           iter = 5000, warmup = 2000, chains = 2, cores = 2, seed = 13,
                                           file = here::here("takeover_performance_cogload_time_budget/Models/mean_decel_skewnormal_reg_priors_maximal"))
# summarising model parameters 
print(summary(mean_decel_skewnormal_reg_priors_maximal), digits = 5)
describe_posterior(mean_decel_skewnormal_reg_priors_maximal , centrality = "mean")

pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "stat_grouped", stat = "mean", group = "n_back")

pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "dens_overlay_grouped", group = "n_back", ndraws = 100)
pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

pp_check(mean_decel_skewnormal_reg_priors_maximal, type = "hist", group = "ttc_criticality_factor")

# model of maximum longitudinal jerk with regularising priors. Random effects structure is maximal
skew_skew_mix = mixture(family = skew_normal(), family = skew_normal(), order = 'mu')

max_long_jerk_skewnormal_mix_reg_priors_maximal <- brm(data = response.vehicle.eye %>%
                                             dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, ho.rt > .15, collision == 0) %>%
                                             dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x)) %>%
                                             dplyr::mutate(ppid = as.factor(ppid)),
                                           family = skew_skew_mix,
                                           prior = c(prior(student_t(3, 0, 2.5), class = "Intercept", dpar = "mu1"),
                                                prior(student_t(3, 0, 2.5), class = "Intercept", dpar = "mu2"),
                                                prior(student_t(3, 0, 2.5), class = "b", dpar = "mu1"),
                                                prior(student_t(3, 0, 2.5), class = "b", dpar = "mu2")),
                                           max_long_jerk ~ n_back * ttc_criticality_factor + (n_back * ttc_criticality_factor | ppid),
                                           control = list(adapt_delta = 0.99999999999),
                                           iter = 5000, warmup = 2000, chains = 2, cores = 2, seed = 13,
                                           file = here::here("takeover_performance_cogload_time_budget/Models/max_long_jerk_skewnormal_mix_reg_priors_maximal"))
# summarising model parameters 
print(summary(max_long_jerk_skewnormal_mix_reg_priors_maximal), digits = 5)
describe_posterior(max_long_jerk_skewnormal_reg_priors_maximal, centrality = "mean")

pp_check(max_long_jerk_skewnormal_mix_reg_priors_maximal, type = "dens_overlay_grouped", group = "n_back", ndraws = 100)
pp_check(max_long_jerk_skewnormal_mix_reg_priors_maximal, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

grand_mean_n_back_dist <- max_long_jerk_skewnormal_reg_priors_maximal %>% 
  epred_draws(newdata = expand_grid(ttc_criticality_factor = c(as.factor(3), as.factor(5)),
                                    n_back = c(TRUE, FALSE)),
              re_formula = NA)

@@

max_long_mixture_6  <- readRDS(here::here("ITS/Work package 6.5 (HiDrive)/TME study/Analysis/max_long_jerk_skewnormal_reg_priors_maximal.rds"))

@@

pp_check(max_long_mixture_6, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
pp_check(max_long_mixture_6, type = "stat_grouped", stat = "mean", group = "n_back.x")

pp_check(max_long_mixture_6, type = "dens_overlay_grouped", group = "n_back.x", ndraws = 100)
pp_check(max_long_mixture_6, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

grand_mean_n_back_dist <- max_long_mixture_6 %>% 
  epred_draws(newdata = expand_grid(ttc_criticality_factor = c(as.factor(3), as.factor(5)),
                                    n_back.x = c(TRUE, FALSE)),
              re_formula = NA)

@@

@@

d <- max_long_mixture_6$data

ggplot(d, mapping = aes(max_long_jerk)) +
  geom_histogram() +
  facet_grid(n_back.x ~ ttc_criticality_factor)
```

@@

```{r}
ggplot(response.vehicle.eye %>%
         dplyr::filter(rt.outlier == FALSE, ho.rt != 0, collision == 0), mapping = aes(x = max_lat)) +
  geom_histogram(aes(fill = as.factor(n_back)), alpha = 0.5, bins = 50, col = "black") +
  facet_grid(n_back ~ttc_criticality.x)

ggplot(response.vehicle.eye %>%
         dplyr::filter(rt.outlier == FALSE, ho.rt != 0, collision == 0), mapping = aes(x = mean_lat_acc)) +
  geom_histogram(aes(fill = as.factor(n_back)), alpha = 0.5, bins = 50, col = "black") +
  facet_grid(n_back ~ttc_criticality.x)
